Automated Verification and Synthesis of Embedded Systems: Advances and Challenges

### 2017-06-19 paper.tex
updated example of inductive invariant
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. For instance, consider the following fragment of C code as shown in Figure~\ref{figure:inductive-invariant}. If we consider the floating-point arithmetic semantics, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. This invariant, however, is not inductive given that $x>0$ before an iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the third iteration. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness. In this specific example, an inductive invariant is $x>1$ since we can check that if $x>1$ before the initial iteration, then also $x>1$ after the \textit{n}-iterations.
%
\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <stdio.h>
#include <assert.h>
#include <math.h>
int main(void) {
 unsigned int i=1, n = -1;
 float x=2;
 while (i<=n) {
   x = ((2*x) - 1);
   if(isinf(x)) {
     printf("\%d\n", i);
     break;
   }
   i++;
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for the \textit{k}-induction algorithm.}
\label{figure:inductive-invariant}
\end{figure}

>>>>

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. For instance, consider the following fragment of C code as shown in Figure~\ref{figure:inductive-invariant}. If we consider the floating-point arithmetic semantics, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. This invariant, however, is not inductive given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness. In this specific example, an inductive invariant is $x>1$ since we can check that if $x>1$ before the initial iteration, then also $x>1$ after the \textit{k}-iterations.
%
\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for the \textit{k}-induction algorithm.}
\label{figure:inductive-invariant}
\end{figure}
